The Concurrent Calculi Formalisation Benchmark

### 2024-02-20 short-paper/short-paper.tex
Fix some minor issues and remove some comments.

\section{The Benchmark and its Design}\label{sec:design-discussion}
\alb{Next paragraph needs rewriting}
In this section, we outline the factors considered in designing the benchmark challenges.
We begin with some general remarks, then describe the individual design considerations for each challenge problem, and the evaluation criteria for solutions.

Similarly to the authors of the POPLMark challenges, we seek to answer several questions:
\begin{enumerate}[label=\textbf{(Q\arabic*)},leftmargin=10mm]
\item\label{item:rq1} What is the current state of the art in tools for mechanising the meta-theory of process calculi?
\item\label{item:rq2} Which techniques and best practices can be recommended when starting new mechanisation projects involving process calculi?
\item\label{item:rq3} Which improvements are needed to make mechanisation tools more user-friendly with regards to the issues faced when mechanising process calculi?
\end{enumerate}
%
\todo{Frederik: I'm not sure this note is necessary}
Note that the challenges revolve around the mechanisation of the
\emph{meta-theory} of process calculi, rather than interpreters or
type checkers for the latter.

@@

\subsection{Scope extrusion}

This challenge revolves around the formalization explicit scope
extrusion, by which a process can send restricted names to another
process, as long as the restriction can safely be ``extruded'' (\ie,
expanded) to include the receiving process.  
The setting for this challenge is a ``classic'' untyped
\( \pi \)-calculus, where (unlike the calculi in the other challenges)
names can be sent and received, and bound by input constructs.  We define two
different semantics for our system:
\begin{enumerate}
\item A reduction system: this avoids explicit reasoning
about scope extrusion via structural congruence, allowing implementors to explore
alternative ways to encode the latter (\eg via context or via congruent refinements)
\item A (early) labelled transition system
\end{enumerate}
The objective of this
challenge is to prove that the two semantics are equivalent up to
structural congruence.

@@

Given those peculiarities, the syntax and semantics has been
mechanized from a an early age (see~\cite{Melham94}) with many proof
assistants and so many encoding styles that it would take too long to
review.  They range from concrete encodings based on named syntax
(Melham's op.cit.) to basic de Bruijn (cite) and locally nameless
representation~\cite{Chargueraud2012} (\eg~\cite{Castro2020}).
Nominal approaches are also common, see~\cite{Bengtson2009}, but they
may be problematic w.r.t.~proof assistants based on dependent types.

@@

Higher-order abstract syntax (HOAS) has seen extensive and successful
use in formal reasoning~\cite{Despeyroux2000,Honsell2001,Tiu2010}. Its
weak form aligns well with mainstream inductive proof assistants,
simplifying the encoding of typing systems and operational
semantics. However, when addressing more intricate concepts like
bisimulation, HOAS necessitates extensions. These extensions may take
the form of additional axioms~\cite{Honsell2001} or require specialized proof assistants
like Abella, which introduces a dedicated quantifier for handling
properties related to names~\cite{GacekMN11}.

>>>>

\section{The Benchmark and its Design}\label{sec:design-discussion}
In this section, we outline the factors considered in designing the benchmark challenges.
We begin with some general remarks, then describe the individual design considerations for each challenge problem, and the evaluation criteria for solutions.

Similarly to the authors of the POPLMark challenges, we seek to answer several questions:
\begin{enumerate}[label=\textbf{(Q\arabic*)},leftmargin=10mm]
\item\label{item:rq1} What is the current state of the art in tools for mechanising the meta-theory of process calculi?
\item\label{item:rq2} Which techniques and best practices can be recommended when starting new mechanisation projects involving process calculi?
\item\label{item:rq3} Which improvements are needed to make mechanisation tools more user-friendly with regards to the issues faced when mechanising process calculi?
\end{enumerate}

@@

\subsection{Scope extrusion}
This challenge revolves around the mechanisation of explicit scope
extrusion, by which a process can send restricted names to another
process, as long as the restriction can safely be ``extruded'' (\ie,
expanded) to include the receiving process.  
The setting for this challenge is a ``classic'' untyped
\( \pi \)-calculus, where (unlike the calculi in the other challenges)
names can be sent and received, and bound by input constructs.  We define two
different semantics for our system:
\begin{enumerate}
\item A reduction system: this avoids explicit reasoning
about scope extrusion via structural congruence, allowing implementors to explore
alternative ways to encode the latter (\eg via context or via congruent refinements)
\item A (early) labelled transition system
\end{enumerate}
The objective of this
challenge is to prove that the two semantics are equivalent up to
structural congruence.

@@

Given those peculiarities, the syntax and semantics has been
mechanised from an early age (see~\cite{Melham1994}) with many proof
assistants and so many encoding styles that it would take too long to
review.  They range from concrete encodings based on named syntax
(Melham's op.cit.) to basic de Bruijn (cite) and locally nameless
representation~\cite{Chargueraud2012} (\eg~\cite{Castro2020}).
Nominal approaches are also common, see~\cite{Bengtson2009}, but they
may be problematic w.r.t.~proof assistants based on dependent types.

@@

Higher-order abstract syntax (HOAS) has seen extensive and successful
use in formal reasoning~\cite{Despeyroux2000,Honsell2001,Tiu2010}. Its
weak form aligns well with mainstream inductive proof assistants,
simplifying the encoding of typing systems and operational
semantics. However, when addressing more intricate concepts like
bisimulation, HOAS necessitates extensions. These extensions may take
the form of additional axioms~\cite{Honsell2001} or require specialised proof assistants
like Abella, which introduces a dedicated quantifier for handling
properties related to names~\cite{GacekMN11}.
